---
title: "p8105_hw2_sz2872"
author: "Shunyi Zhang"
output: github_document
editor_options: 
  chunk_output_type: console
---

This is my solution for hw2.

```{r setup}
library(tidyverse)
```


## Problem 1

Read and Clean the Mr. Trash Wheel sheet.

```{r tidy mr_trashwheel}
mr_trashwheel_df = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                     sheet = "Mr. Trash Wheel",
                     range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) 
head(mr_trashwheel_df)
```

Read and Clean precipitation data for 2017 and 2018.

```{r tidy precipitation for 2017 and 2018}
precip_2017 =
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = "2017 Precipitation",
                     skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = "2017") %>% 
  relocate(year)

precip_2018 =
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = "2018 Precipitation",
                     skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = "2018") %>% 
  relocate(year)
```

Bind precip_17 and precip_18

```{r}
precip_tidy =
  bind_rows(precip_2017, precip_2018) %>% 
  mutate(month = month.name[month]) 
head(precip_tidy)
```

After clean and tidy both data frames, here are the summaries for both datasets.

For `mr_trashwheel_df`, it has `r mr_trashwheel_df %>% nrow()` rows and `r mr_trashwheel_df %>% ncol()` columns.The total number of observations is 4816. The key variables are `r mr_trashwheel_df %>% ls()`.

For `precip_tidy`, it has `r precip_tidy %>% nrow()` rows and `r precip_tidy %>% ncol()` columns.The total number of observations is 72. The key variables are `r precip_tidy %>% ls()`.

The median number of sports balls in a dumpster in 2017 is `r mr_trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.

The total precipitation in 2018 is `r precip_tidy %>% filter(year == 2018) %>% pull(total) %>% sum()`.


## Problem 2

Read and Clean the NYC subway data.

```{r}
transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, entrance_type, vending, ada) %>% 
  mutate(entry = ifelse(entry == "YES",TRUE, FALSE)) %>% 
  mutate_at(vars(route8:route11), as.character) 
head(transit_df)
```

The `transit_df` dataset contains `r transit_df %>% ls()` as variables.

After load the .csv file, the dataset is named `transit_df`. The next step following is unifing the variables' names using `janitor::clean_names`. Then, using the `select` function to pick the variables as order. Moving to the next part, I apply the `mutate` and `ifelse` functions for the purpose of convert the entry variable from character to a logical variable. The last step is converting variables (route8 to route11) from double variables to character variables. 

As all processes are done, the `transit_df` has `r transit_df %>% dim()`. 

I don't think the data is tidy enough. For `route1:route11`, it is reasonable to use `pivot_longer` to make the datast easier to work with. 

#### Answer Qestions

There are `r transit_df %>% distinct(station_name, line) %>% count()` distinct stations presented in the NYC transit dataset.

Among the distinct stations, `r transit_df %>% filter(ada == TRUE) %>% distinct(station_name, line) %>% count()` of them are ADA compliant. 

The proportion of station entrances / exits without vending allow entrance is.
'it should be one of the following, but who knows, I have trouble reading the sentance. Waiting for help. '
`r transit_df %>% filter(entry == TRUE, vending == "NO") %>% nrow() / transit_df %>% filter(entry == TRUE) %>% nrow()`
`r transit_df %>% filter(entry == TRUE) %>% nrow()`
`r transit_df %>% filter(vending == "NO", entry == TRUE) %>% nrow()`


## Problem 3

Clean the Data in pols-month.csv.

```{r clean pols_month}
pols_month = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% #load data
  janitor::clean_names() %>%                                     #clean name
  separate(mon, c("year", "month", "day"), sep = "-") %>%        #separate mon
  mutate(year = as.integer(year), month = as.integer(month)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>%                          #get month name
  mutate(prez_gop = recode(prez_gop, `2` = 1)) %>%               #recode unlisted observations 
  pivot_longer(c("prez_gop", "prez_dem"), 
               names_to = "president",
               names_prefix = "prez_",
               values_to = "gop_dem"
               ) %>%                                             #create 'president'
  filter(gop_dem == 1) %>%                                       #fit 'president'
  relocate(year, month, president) %>%                           #set format
  select(-"day", -"gop_dem")                                     #limit variables
head(pols_month)
```

Clean the Data in snp.csv

```{r}
snp_df =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year")) %>% 
  relocate(year, month) %>% 
  select(-"day") %>% 
  mutate(year = as.integer(year), month = as.integer(month)) %>%
  arrange(year, month) %>% 
  mutate(month = month.name[month]) 
head(snp_df)
```

Clean the Data in unemployment.csv

```{r}
unemp_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(c("Jan":"Dec"),
               names_to = "month",
               values_to = "percentage") %>% 
  mutate(month = month.name[match(month, month.abb)]) %>% 
  janitor::clean_names() %>% 
  mutate(year = as.integer(year))
head(unemp_df)
```

